---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r readData, echo = TRUE}
stepsData <- read.csv("C:/Users/Michael/Documents/dataScience/activity.csv")
stepsData$date <- as.Date(stepsData$date, format="%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r stepsHistogram, echo = TRUE, fig.path="figure/"}
totalSteps <- tapply(stepsData$steps, stepsData$date, sum)

hist(totalSteps)

mean(totalSteps, na.rm = TRUE)

median(totalSteps, na.rm = TRUE)
```

## What is the average daily activity pattern?
Plot of the average number of steps for each time interval and interval with the maximum number of avg. steps.
```{r dailyActivityPattern, echo=TRUE, fig.path="figure/"}
avgStepsByInterval <- aggregate(steps ~ interval, data=stepsData, FUN=mean)
with(avgStepsByInterval, plot(interval, steps, type="l"))
avgStepsByInterval[which.max(avgStepsByInterval$steps), "interval"]
```


## Imputing missing values
Number of missing values & histogram of steps when NA values are replaced with the interval mean:
```{r imputedValues, echo=TRUE, fig.path="figure/"}
sum(is.na(stepsData$steps))

stepsDataModified <- stepsData
stepsDataModified$steps <- as.numeric(stepsDataModified$steps)
naRows <- which(is.na(stepsDataModified$steps), arr.ind = TRUE)
stepsDataModified$steps[naRows] <- avgStepsByInterval[match(stepsDataModified$interval[naRows], avgStepsByInterval$interval), "steps"]
totalStepsModified <- tapply(stepsDataModified$steps, stepsDataModified$date, sum)

hist(totalStepsModified, xlab="steps")

mean(totalStepsModified)

median(totalStepsModified)
```

## Are there differences in activity patterns between weekdays and weekends?
Weekdays have a higer maximum number of steps when compared to weekends:
```{r weekdayPatterns, echo=TRUE, fig.path="figure/"}
weekends <- c("Saturday", "Sunday")
stepsDataModified$weekday <- factor((weekdays(stepsDataModified$date) %in% weekends), levels=c(TRUE,FALSE), labels=c("weekend", "weekday"))

weekdayData <- subset(stepsDataModified, weekday=="weekday")
totalWeekday <- aggregate(steps ~ interval, data=weekdayData, FUN=sum)

weekendData <- subset(stepsDataModified, weekday=="weekend")
totalWeekend <- aggregate(steps ~ interval, data=weekendData, FUN=sum)

par(mfrow=c(2,1))
with(totalWeekday, plot(interval, steps, type="l", main ="weekday"))
with(totalWeekend, plot(interval, steps, type="l", main="weekend"))
```
